---
title: "STA4273 HW 2"
author: "John He"
date: "2022-10-04"
output: html_document
---

# 6.1
$$
\int_0^{\frac{\pi}{3}}sin(t)dt\\
\bigl[-cos(t)\bigr]_0^\frac{\pi}{3}\\
-cos(\frac{\pi}{3})-(-cos(0))\\
-\frac{1}{2}+1\\
\frac{1}{2}
$$
```{r}
m <- 1000000
t <- runif(m,0,pi/3)
theta.hat <- mean(sin(t)) * (pi/3)
print(theta.hat)
```
$$
T_1,...,T_{m}\sim{Unif(0,\frac{\pi}{3})}\\
\hat{\theta}=(\frac{\pi}{3})\frac{1}{m}
\sum_{n=1}^{m}\sin(T_i)
$$

# 6.3
$$
\int_0^{0.5}e^{-x}dx\\
\bigl[-e^{-x}\bigr]_0^{0.5}\\
-e^{-0.5}-(-e^{-0})\\
1-e^{-0.5}\\
\approx0.39346934
$$
```{r}
# Theta Hat
# Sample from Uniform(0,0.5) distribution
theta.hat <- function(m){
  x1 <- runif(m,0,0.5)
  theta.hat <- mean(exp(-x1)) * (0.5)
  print(theta.hat)
}
theta.hat(1000000)
```
$$
X_1,...,X_{m}\sim{Unif(0,0.5)}\\
\hat{\theta}=(\frac{1}{2})\frac{1}{m}
\sum_{i=1}^{m}e^{-X_i}
$$
```{r}
# Theta Star
# Sample from Exponential(1) distribution
theta.star <- function(n){
  x2 <- rexp(n, 1)
  g <- (exp(-x2)/(exp(-x2))*(x2>0)*(x2<0.5))
  theta.star <- mean(g)
  vartstar <- var(g)
  print(theta.star)
}
theta.star(1000000)
```

$$
X_1,...,X_{n}\sim{Exp(1)}\\
\hat{\theta}=\frac{1}{n}
\sum_{j=1}^{n}\frac{e^{-X_j}}{e^{-X_j}}I(0\le X_j\le0.5)\\
\hat{\theta}=\frac{1}{n}
\sum_{j=1}^{n}I(0\le X_j\le0.5)
$$

# 6.4
Monte Carlo estimate of Beta(3,3) cdf
Use function to estimate F(x) for x=0.1,0.2,...,0.9
Compare estimates with values returned by the pbeta function
$$
X_1,...,X_{n}\sim{Beta(3,3)}\\
f(x)=\frac{\Gamma(\alpha+\beta)}{\Gamma(\alpha)\Gamma(\beta)}x^{\alpha-1}(1-x)^{\beta-1}, x\in[0,1]\\
f(x)=\frac{\Gamma(3+3)}{\Gamma(3)\Gamma(3)}x^{3-1}(1-x)^{3-1}\\
f(x)=\frac{\Gamma(6)}{\Gamma(3)\Gamma(3)}x^{2}(1-x)^{2}\\
f(x)=\frac{5!}{2!2!}x^{2}(1-2x+x^{2})\\
f(x)=\frac{5*4*3}{2*1}(x^{2}-2x^{3}+x^{4})\\
f(x)=30(x^{4}-2x^{3}+x^{2})\\
f(x)=30x^{4}-60x^{3}+30x^{2}\\
$$
$$
F(0)=0\\
F(x)=\int(30x^{4}-60x^{3}+30x^{2})dx\\
F(x)=6x^{5}-15x^{4}+10x^{3}+C\\
0=6(0)^{5}-15(0)^{4}+10(0)^{3}+C\\
C=0\\
F(x)=6x^{5}-15x^{4}+10x^{3}\\
$$

$$
\theta=\int_0^{1}(30x^{4}-60x^{3}+30x^{2})dx\\
X_1,...,X_{m}\sim{Unif(0,1)}\\
\hat{\theta}=\frac{1}{m}
\sum_{i=1}^{m}(30X_i^{4}-60X_i^{3}+30X_i^{2})I(0\le X_i\le1)
$$
```{r}
# Monte Carlo estimate of Beta(3,3) cdf
# Estimate F(x) for x=0.1,0.2,...,0.9
beta33estimate <- function(n,lower,upper){
  x3 <- runif(n,0,1)
  g3 <- (30*x3^4-60*x3^3+30*x3^2)*(x3>lower)*(x3<upper)
  beta33estimate <- mean(g3)
}
index <- 0
estimate <- rep(NA, 10)
for (i in seq(0, 1, 0.1)) {
  estimate[index] <- beta33estimate(1000000,0,i)
  index <- index + 1
  # Monte Carlo estimate of Beta(3,3) cdf
  # for x=0.1,0.2,...,0.9
  # and x=1
}
index <- 0
actual <- rep(NA, 10)
for (i in seq(0, 1, 0.1)) {
  actual[index] <- pbeta(i,3,3)
  index <- index + 1
  # Actual values of Beta(3,3) cdf
  # for x=0.1,0.2,...,0.9
  # and x=1
}
x <- c(0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)
df <- data.frame(x,estimate, actual)
print(df)
```

# 6.13
$$
g(x)=\frac{x^2}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}, x>1
$$
# Find two importance functions f1 and f2 that are supported on (1,infinity) and are "close" to g(x)
# Which one should produce the smaller variance in estimating G(x) by importance sampling? Explain.
$$
G(x)=\int_1^{\infty}\frac{x^2}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}dx
$$
# f1 follows a Exponential(0.5) distribution
$$
F1\sim{Exp(\frac{1}{2})}\\
f1(x)=\frac{1}{2}e^{-\frac{x}{2}}, x\ge0\\
$$

$$
G(x)=\int_1^{\infty}\frac{\frac{x^2}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}}{\frac{1}{2}e^{-\frac{x}{2}}}\frac{1}{2}e^{-\frac{x}{2}}dx\\
G(x)=\int_1^{\infty}\frac{2}{\sqrt{2\pi}}x^2e^{\frac{-x^2+x}{2}}\frac{1}{2}e^{-\frac{x}{2}}dx\\
G(x)=\frac{2}{\sqrt{2\pi}}\int_1^{\infty}x^2e^{\frac{-(x^2-x)}{2}}\frac{1}{2}e^{-\frac{x}{2}}dx\\
G(x)=\frac{\sqrt2}{\sqrt{\pi}}\int_1^{\infty}x^2e^{\frac{-(x^2-x+\frac{1}{4})+\frac{1}{4}}{2}}\frac{1}{2}e^{-\frac{x}{2}}dx\\
G(x)=\frac{\sqrt2}{\sqrt{\pi}}e^{\frac{1}{8}}\int_1^{\infty}x^2e^{-\frac{(x-\frac{1}{2})^2}{2}}\frac{1}{2}e^{-\frac{x}{2}}dx\\
X_1,...,X_{m}\sim{Exp(\frac{1}{2})}\\
F1=\frac{\sqrt2}{\sqrt{\pi}}e^{\frac{1}{8}}\frac{1}{m}\sum_{i=1}^{m}X_i^{2}e^{-\frac{(X_i-\frac{1}{2})^2}{2}}I(X_i\ge1)
$$
# f2 follows a Gamma(3,2) distribution
$$
F2\sim{Gamma(3,2)}\\
f2(x)=\frac{1}{2^3\Gamma(3)}x^{3-1}e^{-\frac{x}{2}}, x>0\\
f2(x)=\frac{1}{8*2!}x^{2}e^{-\frac{x}{2}}\\
f2(x)=\frac{1}{16}x^{2}e^{-\frac{x}{2}}\\
$$

$$
G(x)=\int_1^{\infty}\frac{\frac{x^2}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}}{\frac{1}{16}x^{2}e^{-\frac{x}{2}}}\frac{1}{16}x^{2}e^{-\frac{x}{2}}dx\\
G(x)=\int_1^{\infty}\frac{16}{\sqrt{2\pi}}e^{\frac{-x^2+x}{2}}\frac{1}{16}x^{2}e^{-\frac{x}{2}}dx\\
G(x)=\frac{16}{\sqrt{2\pi}}\int_1^{\infty}e^{\frac{-(x^2-x)}{2}}\frac{1}{16}x^{2}e^{-\frac{x}{2}}dx\\
G(x)=\frac{16}{\sqrt{2\pi}}e^{\frac{1}{8}}\int_1^{\infty}e^{-\frac{(x-\frac{1}{2})^2}{2}}\frac{1}{16}x^{2}e^{-\frac{x}{2}}dx\\
X_1,...,X_{m}\sim{Gamma(3,2)}\\
F2=\frac{16}{\sqrt{2\pi}}e^{\frac{1}{8}}\frac{1}{m}\sum_{i=1}^{m}e^{-\frac{(X_i-\frac{1}{2})^2}{2}}I(X_i\ge1)
$$

```{r}
## define the integrated function
G <- function(x) {x^2/(sqrt(2*pi))*exp(-x^2/2)}
## integrate the function from 1 to infinity
integrate(G, lower = 1, upper = Inf)

# Estimate G(x) with F1
# f1~Exp(0.5)
F1 <- function(n){
  x1 <- rexp(n,0.5)
  g1 <- sqrt(2/pi)*exp(1/8)*(x1^2)*exp(-(x1-0.5)^2/2)*(x1>=1)
  F1 <- mean(g1)
  varF1 <- var(g1)
  print(F1)
  print(varF1)
}
F1(1000000)

# Estimate G(x) with F2
# f2~Gamma(3,2)
F2 <- function(n){
  x2 <- rgamma(n,3,,2)
  g2 <- (16/sqrt(2*pi))*exp(1/8)*exp(-(x2-0.5)^2/2)*(x2>=1)
  F2 <- mean(g2)
  varF2 <- var(g2)
  print(F2)
  print(varF2)
}
F2(1000000)
```
# F1 has a variance around 0.254
# F2 has a variance around 1.19
# Using f1 produces a smaller variance

# 6.14
# Obtain a Monte Carlo estimate of G(x) by importance sampling
$$
G(x)=\int_1^{\infty}\frac{x^2}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}dx
$$

$$
X\sim{N(0,1)}\\
f(x)=\frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac{1}{2\sigma^2}(x-\mu)^2}\\
f(x)=\frac{1}{\sqrt{2\pi(1)}}e^{-\frac{1}{2(1)}(x-0)^2}\\
f(x)=\frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}\\
$$

$$
G(x)=\int_1^{\infty}\frac{\frac{x^2}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}}{\frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}}\frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}dx\\
G(x)=\int_1^{\infty}x^2{\frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}}dx\\
X_1,...,X_{m}\sim{N(0,1)}\\
\hat{\theta}=\frac{1}{m}
\sum_{i=1}^{m}X_i^{2}I(X_i\ge1)
$$

```{r}
## define the integrated function
G <- function(x) {x^2/(sqrt(2*pi))*exp(-x^2/2)}
## integrate the function from 1 to infinity
integrate(G, lower = 1, upper = Inf)

# Estimate G(x)
# Sample from Normal(0,1) distribution
Ghat <- function(n){
  x4 <- rnorm(n,0,1)
  g <- (x4^2)*(x4>=1)
  Ghat <- mean(g)
  print(Ghat)
}
Ghat(1000000)
```